### 2025-02-10 paper/conclusions.tex

\section{Conclusions\label{sec:conclusions}}

The paper demonstrates a technique of developing high performance applications with
strong correctness guarantees.  The key insight lies in using a proof assistant
in cooperation with a high-performance language of choice.  This gives a clear
separation of concerns that is very difficult to achieve within a single
language.  The proof assistant is used to design a specification, prove
all the correctness invariants of interest and performs an extraction
into a high-performance language of choice.  This may take a non-trivial
effort, but correctness \emph{is} demanding!

Having a trusted specification as well as entire code-generation pipeline
within a single dependently-typed framework is incredibly powerful.
As we have demonstrated at the example of the neural network, we can introduce domain-specific
optimisations and even swap the backend in case its performance is unsatisfying.
For our example, the entire framework that includes array theory, DSL,
optimisations and extraction is about 2000 lines of Agda code.  We managed
to introduce a new backend in about two days and match performance of
the hand-written code.

A lot of pieces that we have developed in this
paper can be reused in other numerical applications.  We used dependent
types to guarantee the absence of out-of-bound indexing, certain function being
inverses as well as well-scopedness and well-typedness of our DSL.
However, there are many more opportunities that we did not explore.
For example, one can prove the correctness of optimisations, relating
evaluation of optimised an non-optimised expressions.  We can provide
more guarantees when we run extraction, \eg{} we can formalise some
aspects of the backend language and relate them to our DSL.
As for the DSL itself, we can try extending it with internal let
constructions which should improve our C code generation as well as
facilitate optimisations of the derivatives.

There are indeed plenty of opportunities, but the key point is this.
Correctness and performance are competing requirements when it
comes to application design.  Therefore, such a cooperation between
correctness-oriented and performance-oriented tools is likely to
persist.  With this work we demonstrate that such cooperation
is feasible in practice.