### 2025-02-13 paper/relatedwork.tex
Add notes for my work.

\section{Related Work\label{sec:relatedwork}}

% - "Verified Tensor-Program Optimization Via High-level Scheduling Rewrites"
% - "Efficient Differentiable Programming in a Functional Array-Processing Language"
% - "Verified tensor-program optimization via high-level scheduling rewrites"
% - "Indexed Streams: A Formal Intermediate Representation for Fused Contraction Programs"
% - "You only linearize once"
% - Dependent ML
% - ATS
% - Jax
% - Remora