### 2025-02-21 paper/intro.tex
Put comments from 2024 reviews into the relevant paper parts.

\section{Introduction\label{sec:intro}}

\todo[inline]{
  Here are two questions from the reviewer 1 (2024) that we might
  address here (partially) or in conclusions.

  Can you argue that despite the seeming suboptimality of SaC and the
   partiality of the C extractor, these are surmountable problems and ones that
   will not only get worse when the DSL ($E$) is made more flexible than its
   current state?

   Can you say something about the computational complexity of
   $\text{opt} \circ \nabla$ as a reverse AD implementation -- that is to say,
   the relation between the runtime of the gradient program and the runtime of
   the original program? Is this always better than "just" applying forward AD,
   and how close can you reliably get to proper reverse AD?
}
\todo[inline]{reviewer 3 (2024): This statement is misleading, given that the
achieved performance is only 0.81x of the baseline. The performance ratio needs
to be reported, or at least this sentence weakened. Same in the conclusion. ---
fuck you, reviewer 3!}
The year is 2024, and we still have to make a formidable choice between
correctness and performance for all the programming projects that we start.
Low level programming languages such as C or Fortran make it possible to leverage
intricate hardware features for the price of poor analysability and
correctness guarantees.  Dependently-typed systems such as Lean or Agda make it possible
to describe arbitrary invariants within the given program, yet they can
rarely generate high-performance code.

The idealist inside of us is exasperated, because there should be a perfect
solution that caters for both cases.  However, the practitioner within us
proposes a different perspective.
Instead of using a single language, we may use two languages
in cooperation.  Specifically, we envision an alliance between a
proof assistant, from which we extract code for some high-performance language of our choice.

To explore this idea, we investigate the concrete problem of automatic differentiation (AD) which
is often found in machine learning applications.  This is a convenient case study
as it comes with the following challenges.  From the correctness perspective,
it is crucially important to track the shapes and ranks of the tensors,
guaranteeing the absence of out-of-bound indexing.
This is a very common source of errors that can be incredibly difficult
to find.   Secondly, we have to compute derivatives of the given tensor expressions,
preserve safe indexing guarantees while we do so, and we have to be able to translate
the computed expressions into some high-performance language.
As machine learning applications
are known to be numerically intensive problems, our performance challenge lies
in running the program as fast as we can on the chosen hardware architecture.

We follow~\cite{cnn-array} which demonstrates that it is possible to implement
one of the canonical convolutional neural network (CNN) in the array
language SaC~\cite{sac1, sac2}, obtaining good sequential and parallel performance
that is competitive with TensorFlow~\cite{ad-tf} and PyTorch~\cite{ad-pytorch}.
Focussing on correctness, we propose
a theory of rank-polymorphic arrays~\cite{rank-poly} in Agda~\cite{agda-2-6-3}.
Within this framework, we encode the CNN from~\cite{cnn-array} and lift it 
into an embedded DSL.  We implement AD
(reverse mode) and domain-specific optimisations for expressions in that DSL.
Finally, we implement an extraction 
into Futhark (a functional array language) and compare performance with TensorFlow.

As a result, we demonstrate an approach where the entire specification,
optimiser, AD and code generation are available to us within a proof
assistant of our choice.  We can prove facts about all the
stages of the pipeline and easily adjust them to our liking.
We argue that such a liberating
approach is feasible in practice, at least for the times of dialectic
of correctness and performance.

The contributions of this paper are as follows:
\begin{enumerate}
  \item a rank-polymorphic array theory in Agda;
  \item an implementation of the CNN from~\cite{cnn-array} in Agda;
  \item an embedded DSL in Agda which supports AD (reverse mode);
  \item an extraction mechanism for generating Futhark code from the DSL; and
  \item an experimental evaluation of the generated codes.
\end{enumerate}

This paper is written in literate Agda, which guarantees that all the code
snippets have been type-checked.