### 2025-02-27 paper/conclusions.tex
Pass throught the paper.

\section{Conclusions\label{sec:conclusions}}

The paper demonstrates a technique of developing high performance
applications with strong correctness guarantees, including the absence
of out-of-bound indexing, certain functions being inverses,
well-scopedness and well-typedness of the embedded DSL and
semantics-preserving optimisations.

The key insight lies in using a proof assistant in cooperation with a
high-performance language of choice. This gives a clear separation of
concerns that is very difficult to achieve within a single language.
The proof assistant is used to design a specification, prove all the
correctness invariants of interest and performs an extraction into a
high-performance language, in our case Futhark, although only a
relatively small part of our work is specific to that language.

Having a trusted specification as well as entire code-generation
pipeline within a single dependently-typed framework is incredibly
powerful. As we have demonstrated at the example of the neural
network, we can introduce domain-specific optimisations and
transformations, such as automatic differentiation. For our example,
the entire framework that includes array theory, DSL, optimisations
and extraction is about 2000 lines of Agda code.

A lot of pieces that we have developed in this paper can be reused in
other numerical applications. However, there are many more
opportunities that we did not yet explore. We can formalise other
correctness criteria for our specification spanning from verified
extraction to functional correctness of the actual application such
as CNN robustness.
Choosing DSL primitives so that specificational clarity and
performance requirements are at balance is an interesting question.
Guaranteeing asymptotic efficiency of the AD implementation is another
interesting research direction.  Currently we rely on
optimisations which is known to be a fragile technique, despite its efficiency 
in many cases.  Identifying a normal form for the optimisations would
be a more fundamental approach.  Richer languages with more
control flow may require us choosing more powerful
encoding for the adjoint structures.

There are indeed plenty of opportunities, but the key point is this.
Correctness and performance are competing requirements when it comes
to application design. Therefore, such a cooperation between
correctness-oriented and performance-oriented tools is likely to
persist. With this work we demonstrate that such cooperation can be
done using fairly straightforward means, by Agda standards, and obtain
compelling practical performance.